### Working on tadpoles dataset

```{r}
library(ggplot2)
library(ggthemes)
library(tidyverse)
library(tidybayes)
library(purrr)
library(rethinking)

data(reedfrogs)
d <- reedfrogs

rm(reedfrogs)
detach(package:rethinking, unload = T)
library(brms)

# Making tank the cluster variable
d <- d %>% mutate(tank=1:nrow(d)) %>% as_tibble()
head(d)
```

### Simple fixed effects model

Each tank has its own survival probability, there is no adaptive pooling of information across the clusters

```{r}
d$tank <- as.factor(d$tank)

b12.1 <- brm( data = d, family = 'binomial',
              surv | trials(density) ~ 0 + tank,
              prior = c(prior(normal(0,5), class=b)),
              iter = 2000, warmup = 500, chains = 4, cores = 4, seed = 12)

summary(b12.1)
# plot(b12.1)
```

Drawing posterior samples

```{r}
post <- posterior_samples(b12.1)
post %>%
  select(-lp__) %>%
  gather() %>%
  group_by(key) %>%
  ggplot(aes(x=value, y=reorder(key,value))) +
  stat_pointintervalh(point_interval = mode_hdi, .width = .95,
  size = 3/4, color = "firebrick4") +
  labs(title = "Tadpole survival rate",
  x = NULL, y = NULL) +
  theme_bw() +
  theme(panel.grid = element_blank(),
  panel.grid.major.y = element_line(color = alpha("firebrick4", 1/4), linetype = 3),
  axis.text.y = element_text(hjust = 0),
  axis.ticks.y = element_blank())
```

What does the shrinkage look like the fixed effects based model
```{r}
post_mdn_fx <-
  fixef(b12.1) %>%
  as_tibble() %>%
  select(Estimate) %>%
  bind_cols(d) %>%
  mutate(post_mdn = inv_logit_scaled(Estimate))

post_mdn_fx %>%
  ggplot(aes(x = as.integer(tank))) +
  # geom_hline(yintercept = median(post_mdn)), linetype = 2, size = 1/4) +
  geom_vline(xintercept = c(16.5, 32.5), size = 1/4) +
  geom_point(aes(y = propsurv), color = "orange2") +
  geom_point(aes(y = post_mdn), shape = 1) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_x_continuous(breaks = c(1, 16, 32, 48)) +
  labs(
    title = "Multilevel shrinkage!",
    subtitle = "The empirical proportions are in orange while the model-\nimplied proportions are the black") + 
  annotate("text", x = c(8, 16 + 8, 32 + 8), y = 0,
  label = c("small tanks", "medium tanks", "large tanks")) +
  theme_fivethirtyeight() +
  theme(panel.grid = element_blank())

```

### Simple multi-level model

Adaptively pooling information across all tanks

```{r}

b12.2 <- brm( data = d, family = 'binomial',
              surv | trials(density) ~ 1 + (1|tank),
              prior = c(prior(normal(0,5), class=Intercept),
                        prior(cauchy(0,1), class=sd)),
              iter = 2000, warmup = 500, chains = 4, cores = 4, seed = 12)

summary(b12.2)
```

Simulting the information learned and observing shrinkage

```{r}

post <- posterior_samples(b12.2, add_chain = T)
post_mdn <- coef(b12.2, robust=T)$tank[, ,] %>% 
              as_tibble() %>% 
              bind_cols(d) %>% 
              mutate(post_mdn=inv_logit_scaled(Estimate)) %>%
              bind_cols(post_mdn_fx)

post_mdn

post_mdn %>%
  ggplot(aes(x=as.integer(tank))) +
  geom_hline(yintercept = inv_logit_scaled(median(post$b_Intercept)), linetype=2, size=1/4) +
  geom_vline(xintercept = c(16.5, 32.5), size=1/4) +
  geom_point(aes(y=propsurv), color='orange2') +
  geom_point(aes(y=post_mdn1), color='royalblue') +
  geom_point(aes(y=post_mdn), shape=1) +
  coord_cartesian(ylim = c(0,1)) +
  scale_x_continuous(breaks=c(1,16,32,48)) +
  labs(title='Multilevel Shrinkage!',
       subtitle = 'The empirical proportions are in orange while the model-\nimplied proportions are the black circles. The dashed line is\nthe model-implied average survival proportion.\nOpenCircle - randef\nBlue - fixef\nOrange - data\n') +
  theme_fivethirtyeight() +
  theme(panel.grid = element_blank())

```

```{r}

set.seed(12)

ggplot(data = post, 
       aes(x = rnorm(n    = nrow(post), 
                     mean = b_Intercept, 
                     sd   = sd_tank__Intercept) %>% 
             inv_logit_scaled())) +
  geom_density(size = 0, fill = "orange2") +
  scale_y_continuous(NULL, breaks = NULL) +
  ggtitle("Probability of survival") +
  theme_fivethirtyeight()

```

```{r}

```

