---
title: "Chapter 04"
author: "Srivatsa Srinath"
date: "10/04/2019"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE)
library(dplyr);
library(tidyr);
library(rstan);
library(ggplot2);
library(ggthemes);
library(rethinking);
library(HDInterval)
library(gridExtra)

theme_set(theme_tufte(base_family = 'sans'))

```

# Chapter 4

## 4.3 A gaussian model fo height

```{r read_data}

data("Howell1", package = 'rethinking')
d <- Howell1; rm(Howell1)

d2 <- d %>% filter(age>=18)

```

## 4.3.2 The model

```{stan output.var="m04_1"}

data {
  int<lower=1> N;
  vector[N] height;
}
parameters {
  real mu;
  real<lower=0, upper=50> sigma;
}
model {
  target += normal_lpdf(height | mu, sigma);
  target += normal_lpdf(mu | 178, 20);
}

```

Re-organizing the input in a list format as needed by stan

```{r}

dat <- list(N = nrow(d2), height=d2$height)

```

Sampling from the model

```{r}
fit_04_1 <- sampling(m04_1, data = dat, iter = 1000, chains = 4, cores = 4)

```

```{r}

print(fit_04_1, probs = c(0.1, 0.9))

```

Effect of a narrower prior

```{stan output.var="m04_2"}

data {
  int<lower=1> N;
  vector[N] height;
}
parameters {
  real mu;
  real<lower=0,upper=50> sigma;
}
model {
  target += normal_lpdf(height | mu, sigma);
  target += normal_lpdf(mu | 178, 0.1);
}

```

```{r}

dat <- list(N = nrow(d2), height=d2$height)

fit_04_2 <- sampling(m04_2, data = dat, iter = 1000, chains = 4, cores = 4)
```

```{r}

as.data.frame(fit_04_2) %>% select(mu, sigma) %>% cor()

as.data.frame(fit_04_2) %>% select(mu, sigma) %>% summarise_all(mean)

```

Adding a predictor variable; weight

```{stan, output.var="m04_3"}

data {
  int<lower=0> N;
  vector[N] height;
  vector[N] weight;
}
parameters {
  real alpha;
  real beta;
  real<lower=0, upper=50> sigma;
}
model {
  vector[N] mu = alpha + beta*weight;
  height ~ normal(mu, sigma);
  alpha ~ normal(178,20);
  beta ~ normal(0,10);
  //target += normal_lpdf(height|mu, sigma);
  //target += normal_lpdf(mu|178,20);
  //target += normal_lpdf(beta|0,10);
  

}
```


```{stan, output.var="m04_4"}

data {
  int <lower=0> N;
  vector[N] weight;
  vector[N] height;
}
parameters {
  real alpha;
  real beta;
  real<lower=0> sigma;
}
model {
  vector[N] mu = alpha+beta*weight;
  height ~ normal(mu, sigma);
}
```


```{r}


dat <- list(N = nrow(d2), height=d2$height, weight=d2$weight)

fit_04_3 <- sampling(m04_3, data = dat, iter = 1000, chains = 4, cores = 4)

```

Plotting the posterior inference along the original data

```{r}
post <- as.data.frame(fit_04_3)

ggplot(d2, aes(x=weight, y=height)) + 
  geom_point(shape=1, color='dodgerblue') +
  geom_abline(intercept = mean(post$alpha), slope = mean(post$beta))

```

Consider height at weight of 50 only

```{r}
mu_at_50 <- post$alpha + post$beta*50

ggplot() + geom_density(aes(x=mu_at_50), fill='lightskyblue1') +
  labs(x='mu| weight=50')

```

Computing the credible intervals @80

```{r}
HDInterval::hdi(mu_at_50, credMass=0.8)[1:2]

```

Plotting the density at each weight

```{r}

f_mu <- function(x) post$alpha + post$beta*x;
weight_new <- seq(25, 75)
mu <- sapply(weight_new, f_mu) %>%
        as_tibble() %>%
        rename_all(function(x) weight_new) %>%
        mutate(Iter=row_number()) %>%
        gather(weight, height, -Iter) %>%
        group_by(weight) %>%
        mutate(hpdi_l=HDInterval::hdi(height, credMass = 0.8)[1],
               hpdi_r=HDInterval::hdi(height, credMass = 0.8)[2],
               mu=mean(height)) %>%
        mutate(mu=mean(height)) %>%
        ungroup() %>%
        mutate(weight=as.integer(weight))
mu

```


```{r, fig.width=12}
p1 <- ggplot() + geom_point(data=mu%>%filter(Iter<150), aes(x=weight, y=height), alpha=0.05)
p2 <- ggplot() + geom_point(data=d2, aes(x=weight, y=height), shape=1, color='dodgerblue') +
        geom_ribbon(data=mu, aes(x=weight, ymin=hpdi_l, ymax=hpdi_r), alpha=0.1) +
        geom_abline(data=post, intercept = mean(alpha), slope = mean(beta))
p3 <- ggplot() + geom_point(data=mu, aes(x=weight, y=height), alpha=0.05)
grid.arrange(p1, p2, p3, nrow=1)  
```


Prediction Intervals

```{r}

sim_ht <- sapply(weight_new, function(x) rnorm(nrow(post), post$alpha+post$beta*x, post$sigma)) %>% 
            as_tibble() %>%
            rename_all(function(x) weight_new) %>%
            mutate(Iter=row_number()) %>%
            gather(weight, height, -Iter) %>%
            group_by(weight) %>%
            mutate(pi_l=PI(height, prob = 0.8)[1],
                   pi_r=PI(height, prob = 0.8)[2]) %>%
            ungroup() %>%
            mutate(weight=as.integer(weight))
  
sim_ht

```

```{r, fig.width=12}
p1 <- ggplot() + geom_point(data=mu%>%filter(Iter<150), aes(x=weight, y=height), alpha=0.05) +
        labs(subtitle = 'Density at each weight')
p2 <- ggplot() + geom_point(data=d2, aes(x=weight, y=height), shape=1, color='dodgerblue') +
        geom_ribbon(data=sim_ht, aes(x=weight, ymin=pi_l, ymax=pi_r), alpha=0.1) +
        geom_abline(data=post, intercept = mean(alpha), slope = mean(beta)) +
        labs(subtitle = 'Prediction Intervals at 0.95')
p3 <- ggplot() + geom_point(data=d2, aes(x=weight, y=height), shape=1, color='dodgerblue') +
        geom_ribbon(data=mu, aes(x=weight, ymin=hpdi_l, ymax=hpdi_r), alpha=0.1) +
        geom_abline(data=post, intercept = mean(alpha), slope = mean(beta)) +
        labs(subtitle = 'HPDI Intervals at 0.95')
grid.arrange(p1, p2, p3, nrow=1)  

```

