```{r load_packages}
require(loo)
require(rethinking)
require(tidyverse)
require(dplyr)
require(ggplot2)
require(tidyr)
require(tidybayes)
require(bayesplot)
require(brms)

options(mc.cores = parallel::detectCores())
```

3 gaussian prior

```{r}

tibble(
  d = seq(from = -3, to = 3, length.out = 1e4),
  sig_1 = dnorm(x = d, mean = 0, sd = 1),
  sig_0p5 = dnorm(x = d, mean = 0, sd = 0.5),
  sig_0p2 = dnorm(x = d, mean = 0, sd = 0.2),
) %>%
  gather(prior_type, values, -d) %>%
  ggplot(aes(x = d, y = values, col = prior_type)) +
    geom_line(stat = "identity")

tibble(
  sig_1 = rnorm(n = 1e4, mean = 0, sd = 1),
  sig_0p5 = rnorm(n = 1e4, mean = 0, sd = 0.5),
  sig_0p2 = rnorm(n = 1e4, mean = 0, sd = 0.2),
) %>%
  gather(prior_type, values) %>%
  ggplot(aes(x = values, col = prior_type)) + geom_density()

```


Computing WAIC
```{r}
data(cars)

priors <- c(brms::prior(normal(0, 100), class = Intercept),
            brms::prior(normal(0, 10), class = b),
            brms::prior(cauchy(0, 2), class = sigma))

m <- brm(data = cars, family = gaussian,
         formula = dist ~ 1 + speed,
         prior = priors,
         sample_prior = T, iter = 2000, warmup = 500,
         cores = 4, chains = 4)

```

Stan diagnostics

```{r}

mcmc_dens(m)
mcmc_trace(m)

```

Extracting the prior and posterior sample

```{r}

post_s <- posterior_samples(m)
prior_s <- prior_samples(m)

speed_seq <- seq(from = min(cars$speed),
                 to = max(cars$speed),
                 length.out = nrow(cars))

mu_preds <- sapply(speed_seq,
                   function(x) post_s$b_Intercept + post_s$b_speed * x)

dist_preds <- sapply(speed_seq,
                     function(x) rnorm(n = 1e4,
                                       mean = post_s$b_Intercept + post_s$b_speed * x,
                                       sd = post_s$sigma))

dist_vals <- sapply(seq_len(ncol(dist_preds)),
                    function(s) {
                      rethinking::PI(dist_preds[, s])
                    }) %>%
              t %>%
              as_tibble() %>%
              mutate(speed = speed_seq) %>%
              cbind(colMeans(dist_preds)) %>%
              rename(dist_mean = "colMeans(dist_preds)",
                     dist_ll = "5%",
                     dist_ul = "94%")

mu_vals <- sapply(seq_len(ncol(mu_preds)),
                  function(s) {
                    rethinking::PI(mu_preds[, s])
                  }
            ) %>%
            t %>%
            as_tibble() %>%
            mutate(speed = speed_seq) %>%
            cbind(colMeans(mu_preds)) %>%
            cbind(cars$dist) %>%
            rename(dist = "cars$dist",
                   mu_mean = "colMeans(mu_preds)",
                   mu_ll = "5%",
                   mu_ul = "94%")

mu_vals %>%
  ggplot(aes(x = speed)) +
    geom_ribbon(data = dist_vals,
                aes(x = speed, ymin = dist_ll, ymax = dist_ul),
                fill = "grey83") +
    geom_smooth(data = mu_vals,
                aes(y = mu_mean, ymin = mu_ll, ymax = mu_ul),
                fill = "grey70",
                stat = "identity") +
    geom_point(aes(y = dist))

```

Log likelihood computation

In BRMS, Each occasion / data point gets a column and each HMC chain iteration gets a row.
```{r}

f_ll <- log_lik((m))

c_ll <- sapply(seq_len(nrow(post_s)), function(s_idx) {
  mu <- post_s$b_Intercept[s_idx] + post_s$b_speed[s_idx] * cars$speed
  dnorm(x = cars$dist, mean = mu, sd = post_s$sigma[s_idx], log = T)
}) %>% t

sum(f_ll - c_ll)

```


Deviance measurement

```{r}

f_ll %>%
  as_tibble() %>%
  mutate(sums = rowSums(.),
         deviance =  -2 * sums) %>%
  ggplot() +
    geom_density(aes(x = deviance)) +
    labs(title = "Deviance Distribution")
```

The log-pointwise-predictive-density is the total across observations of the logarithm of the average likelihood of each observation.
For pWAIC compute the variance in log-probabilities for each observation i, and then sum up these variances to get the total penalty.

```{r}

ll <- m %>%
      log_lik() %>%
      as_tibble()

df_means <- ll %>%
              exp() %>%
              summarise_all(mean) %>%
              gather(key, means) %>%
              select(means)

df_vars <- ll %>%
              summarise_all(var) %>%
              gather(key, vars) %>%
              select(vars)

avg_ll_p <- df_means %>% log %>% sum
avg_ll_p
p_waic <- df_vars %>% sum
p_waic

waic <- -2 * (avg_ll_p - p_waic)
waic

waic(m)
```

