## Loading the libraries and dataset

```{r}
library(magrittr)
library(tidybayes)
library(rethinking)
data(WaffleDivorce)
d <- WaffleDivorce

rm(WaffleDivorce)
detach(package:rethinking, unload = T)
library(brms)
library(tidyverse)
library(ggrepel)
```
```{r}

str(d)

```

Plotting waffles houses per million of population and divorce rate


```{r}

d %>%
  ggplot(aes(x=WaffleHouses/Population, y=Divorce)) + 
    geom_point(size=1.5, color='firebrick4', alpha=1/2) +
    stat_smooth(method = 'lm', color='firebrick4', fullrange = T, size=1/2, alpha=1/5, fill='firebrick4') +
    geom_text_repel(
      data = d %>% filter(Loc %in% c("ME", "OK", "AR", "AL", "GA", "SC", "NJ")),
      aes(label = Loc),
      size = 3, seed = 1042) +
  ylab('Divorce Rate') +
  theme_bw() +
  theme(panel.grid = element_blank())


```

Standardize the predictor

```{r}

d <- d %>% mutate(MedianAgeMarriage_s=(MedianAgeMarriage - mean(MedianAgeMarriage))/sd(MedianAgeMarriage))

```

Fitting a brms model

```{r}
b5.1 <- brm(
  data = d, family = gaussian,
  formula = Divorce ~ 1 + MedianAgeMarriage_s,
  prior = c(prior(normal(10,10), class = Intercept),
            prior(normal(0,1), class = b),
            prior(uniform(0,10), class = sigma)),
  iter = 2000, warmup = 500, chains = 4, cores = 4, seed = 5
)

print(b5.1)

```

Building predictions using the model

```{r}
# dataset for prediction
nd <- tibble(MedianAgeMarriage_s=seq(from=-3, to=3.5, length.out = 30))

# using fitted 
fitted_data <- fitted(b5.1, newdata = nd) %>% as_tibble() %>% bind_cols(nd)
head(fitted_data)
head(d)

ggplot(fitted_data, aes(x=MedianAgeMarriage_s, y=Estimate)) +
  geom_smooth(stat = 'identity', aes(ymin=Q2.5, ymax=Q97.5)) +
  geom_point(data = d, aes(y=Divorce), shape=1, size=2, color='firebrick4') +
  theme_bw() + theme(panel.grid = element_blank())

```

A model with mean marraige rate

```{r}

d <- d %>% mutate(Marriage_s=(Marriage-mean(Marriage))/ sd(Marriage))

b5.2 <- brm(
  data = d, family=gaussian,
  formula =  Divorce ~ 1 + Marriage_s,
  prior = c(prior(normal(10,10), class = Intercept),
            prior(normal(0,0.5), class = b, coef=Marriage_s),
            prior(normal(0,10), class=sigma)),
  iter = 41000, warmup = 40000, cores = 4, chains = 4, seed = 5
)

print(b5.2)

plot(b5.2)

```

Visualization of the results
```{r}

# Prediction dataframe
nd <- tibble(Marriage_s=seq(-3,3,length.out = 100))

# fitted data
fitted_data <- fitted(b5.2, newdata = nd) %>% as_tibble() %>% bind_cols(nd)

# visualization
d %>%
  ggplot(aes(x=Marriage_s, y=Divorce)) +
  geom_smooth(data = fitted_data, aes(y=Estimate, ymin=Q2.5, ymax=Q97.5), 
              stat='identity', fill = "firebrick", color = "firebrick4", alpha = 1/5, size = 1/4) +
  geom_point(data = d, aes(y=Divorce), size=2, color='firebrick4') +
  coord_cartesian(xlim = range(d$Marriage_s),
  ylim = range(d$Divorce)) +
  ylab("Divorce") +
  theme_bw() +
  theme(panel.grid = element_blank())
```

Fitting a multivariate model

```{r}

b5.3 <- brm(
  data = d, family = gaussian,
  formula = Divorce ~ 1 + Marriage_s + MedianAgeMarriage_s,
  prior = c(prior(normal(10,10), class = Intercept),
            prior(normal(0,0.5), class = b, coef=Marriage_s),
            prior(normal(0,0.5), class = b, coef=MedianAgeMarriage_s),
            prior(uniform(0,10), class = sigma)),
  iter = 41000, warmup = 40000, chains = 4, cores = 4, seed = 5
)

print(b5.3)
plot(b5.3)


```
Using bayesplot to plot the intervals

```{r}

post <- posterior_samples(b5.3)
post %>%
  select(-lp__) %>%
  gather() %>%
  group_by(key) %>%
  ggplot(aes(x=value, y=reorder(key,value))) +
  stat_pointintervalh(point_interval = mode_hdi, .width = .95,
  size = 3/4, color = "firebrick4") +
  labs(title = "My tidybayes-based coefficient plot",
  x = NULL, y = NULL) +
  theme_bw() +
  theme(panel.grid = element_blank(),
  panel.grid.major.y = element_line(color = alpha("firebrick4", 1/4), linetype = 3),
  axis.text.y = element_text(hjust = 0),
  axis.ticks.y = element_blank())

```

# Handling Categorical variables
```{r}

library(rethinking)
data(Howell1)
d <- Howell1

rm(Howell1)
detach(package:rethinking, unload = T)
library(brms)
```

```{r}

glimpse(d)

```

```{r}

b5.15 <- brm(
  data = d, family=gaussian,
  formula = height ~ 1 + male,
  prior = c(prior(normal(178,10), class = Intercept),
            prior(normal(0,0.5), class = b),
            prior(cauchy(0,2), class = sigma)),
  iter = 2000, warmup = 500, chains = 4, cores = 4, seed = 5
)

print(b5.15)
plot(b5.15)
```

Model based predictions
```{r}

post <- posterior_samples(b5.15) %>% as_tibble()
post %>%
  transmute(male_height=b_Intercept+b_male) %>% mean_qi(.width = .89)

nd <- tibble(male=1)
fitted(b5.15,
       newdata = nd,
       summary = F) %>%
  as_tibble() %>%
  ggplot(aes(x = V1, y = 0)) + 
  geom_halfeyeh(fill = "firebrick4",
                point_interval = median_qi, .width = .95) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(subtitle = "Model-implied male heights",
       x = expression(alpha + beta["male"])) +
  theme_bw() +
  theme(panel.grid = element_blank())

```

Many more categories

```{r}

library(rethinking)
data(milk)
d <- milk

rm(milk)
detach(package:rethinking, unload = T)
library(brms)

```


```{r}

str(d)

```

```{r}

b5.16 <- brm(
  data = d, family = gaussian,
  formula = kcal.per.g ~ 0 + clade,
  prior = c(prior(normal(0.5,0.5), class = b),
            prior(cauchy(0,2), class = sigma)),
  iter = 2000, warmup = 500, chains = 4, cores = 4, seed = 5
)

print(b5.16)
plot(b5.16)

```

```{r}

posterior_samples(b5.16, summary=F) %>% select(-lp__)

```

Adding another predictor variable

```{r}

unique(d$clade)

b5.17 <- brm(
  data = d, family = gaussian,
  formula = kcal.per.g ~ 0 + clade + perc.fat,
  prior = c(
            prior(normal(0,.5), class = b),
            prior(normal(0,0.5), class = b, coef=perc.fat),
            prior(cauchy(0,2), class = sigma)),
  iter = 2000, warmup = 500, chains = 4, cores = 4, seed = 5
)

print(b5.17)
plot(b5.17)

```

```{r}

nd <- tibble(
  clade=c('Strepsirrhine'),
  perc.fat=c(0.5)
)

posterior_summary(b5.17)
fitted(b5.17, newdata = nd)
predict(b5.17, newdata = nd)

```
