
```{r}
library(brms)
library(ggplot2)
library(ggthemes)
library(tidyverse)
library(tidybayes)
library(purrr)

```

## Modeling Heights
###Reading the dataset

```{r}

library(rethinking)
data(Howell1)
d <- Howell1

rm(Howell1)
detach(package:rethinking, unload = T)
library(brms)
d2 <- d %>% filter(age >=18)

```

### Simple model of heights - No predictor variable

Fitting a normal distribution of heights

```{r model_fitting}

b4.1 <- brm(data = d2,
            family = gaussian,
            formula = height ~ 1,
            prior = c(prior(normal(178,20), class= Intercept),
                      prior(cauchy(0,1), class = sigma)),
            iter=2000, warmup = 1000, chains = 4, cores = 4, seed = 4)

print(b4.1)
plot(b4.1)
```

Sampling from the fit model
```{r}
post <- posterior_samples(b4.1)

post %>%
  select(b_Intercept:sigma) %>%
  cov()

post %>%
  select(b_Intercept:sigma) %>%
  cov() %>%
  diag()

# extracting info from the posterior samples
head(post)

post %>%
  select(-lp__) %>%
  gather(parameter) %>%
  group_by(parameter) %>%
  summarise(mean=mean(value),
            SD=sd(value),
            `2.5_percentile`= quantile(value, 0.025),
            `97.5_percentile`= quantile(value, 0.975)) %>%
  mutate_if(is.numeric, round, digits=2)

post %>%
  select(-lp__) %>%
  gather(parameter) %>%
  group_by(parameter) %>%
  mean_qi(value)
```

### Plotting
Plotting sigma distribution from the simple fit above
```{r}
post %>%
  select(sigma) %>%
  ggplot(aes(x=sigma)) + geom_density(fill='black') +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(expression(sigma)) +
  theme(panel.grid = element_blank())

post %>%
  ggplot(aes(x=b_Intercept)) + geom_density(fill='black') +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(expression(b_Intercept)) +
  theme(panel.grid = element_blank())

```
### Adding a predictor variable

Visualizing the data

```{r}

ggplot(data = d2, aes(x=weight, y= height)) + 
  geom_point(shape=1, size=2) + 
  theme_bw() +
  theme(panel.grid = element_blank())

```


Model opt
```{r}

b4.3 <- brm(
  data = d2, family = gaussian,
  formula = height ~ 1 + weight,
  prior = c(prior(normal(156,20), class = Intercept),
            prior(normal(0,10), class = b),
            prior(uniform(0,50), class = sigma)),
  chains = 4, cores = 4, iter = 41000, warmup = 40000, seed = 5
)

print(b4.3)
plot(b4.3)

```

Model Summary
```{r}

posterior_summary(b4.3)

posterior_samples(b4.3) %>%
  select(-lp__) %>%
  cor() %>%
  round(digits = 2)

```
The slope and the intercept values seem to be highly correlated.  Centering the weight column seems to reduce this, need to understand why?

```{r}

d2 <- d2 %>% mutate(weight_c=weight - mean(weight))

b4.4 <- brm(
  data = d2, family = gaussian,
  formula = height ~ 1 + weight_c,
  prior = c(prior(normal(156,20), class = Intercept),
            prior(normal(0,10), class = b),
            prior(uniform(0,50), class = sigma)),
  chains = 4, cores = 4, iter = 41000, warmup = 40000, seed = 5
)

print(b4.4)
plot(b4.4)

```

Model summary checking the correlation among the parameters

```{r}

posterior_summary(b4.4)

posterior_samples(b4.4) %>%
  select(-lp__) %>%
  cor() %>%
  round(digits = 2)

```

PLotting the posterior inference

```{r}

num_lines <- 100

d2 %>%
  ggplot(aes(x=weight_c, y=height)) + geom_point(shape=1,size=2, color='royalblue') +
  geom_abline(intercept = fixef(b4.4)[1],
              slope = fixef(b4.4)[2]) +
  geom_abline(intercept = posterior_samples(b4.4)$b_Intercept[1:num_lines],
              slope = posterior_samples(b4.4)$b_weight_c[1:num_lines],
              size=1/3, alpha=0.1) +  
  theme_bw() +
  theme(panel.grid = element_blank())

```

mu_at_50 plot
```{r}

posterior_samples(b4.3) %>% 
  mutate(mu_at_50=b_Intercept+b_weight*50) %>%
  ggplot(aes(x=mu_at_50)) + geom_density(size=0, fill='royalblue') +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(x = expression(mu["height | weight = 50"])) +
  theme_classic() + stat_pointintervalh(aes(y = 0), point_interval = mode_hdi, .width = .95)


```

### predict using the model
```{r}

weight_seq <- tibble(weight=seq(from=25, to=70, by=1))
mu_fitted <- fitted(b4.3, summary = F, newdata = weight_seq) %>% as_tibble() %>% set_names(25:70) %>% mutate(iter=1:n())
mu_fitted <- mu_fitted  %>% gather(weight, height, -iter) %>% mutate(weight=as.numeric(weight))

d2 %>%
  ggplot(aes(x=weight, y=height)) +
  geom_point(data = mu_fitted %>% filter(iter<101), alpha=0.1)

```

Plotting the intervals

```{r}

mu_summary <- fitted(b4.3, newdata = weight_seq) %>% as_tibble() %>% bind_cols(weight_seq)
head(mu_summary)

d2 %>%
  ggplot(aes(x=weight, y=height)) +
  geom_smooth(data = mu_summary, aes(y=Estimate, ymin=Q2.5, ymax=Q97.5),
              stat='identity', fill='grey70', color='black', alpha=1, size=0.5) +
  geom_point(color='navyblue', shape=1, size=2, aplha=2/3) +
  coord_cartesian(xlim=range(d2$weight)) +
  theme(text = element_text(family = "Times"), panel.grid = element_blank())

```

Plotting the actual height values
```{r}

pred_height <- predict(b4.3, newdata = weight_seq) %>% as_tibble() %>% bind_cols(weight_seq)
pred_height

d2 %>%
  ggplot(aes(x=weight, y=height)) +
  geom_ribbon(data=pred_height, aes(y=Estimate, ymin=Q2.5, ymax=Q97.5), fill='grey83') +
  geom_smooth(data=mu_summary, aes(y=Estimate, ymin=Q2.5, ymax=Q97.5), color='black', aplha=1, size=1/2,
              stat = 'identity', fill='grey70') +
  geom_point(aes(y=height), shape=1, size=2, color='navyblue') +
  coord_cartesian(xlim = range(d2$weight),
  ylim = range(d2$height)) +
  theme(text = element_text(family = "Times"),
  panel.grid = element_blank())

```
